---
title: "2011_Medicare_Cleanup"
output:
  html_document: default
  word_document: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

### The first objective is to clean up the 2011 Medicare hospital Provider file for analysis.
* provider_sepDRG

### The second objective is to clean up the National and State files.
* National_Average_CMS
* State_Average_CMS

### The third objective is to merge the National file with the Provider file.
* provider_sepDRG_Nat

### 1. Clean Provider File


##### Load package 
```{r eval = TRUE}
library(tidyverse)
```

##### Load provider data 

```{r eval = TRUE}

# This file contains data at the individual hopsital (provider) level.

provider <- read_csv("C:\\Users\\bliss\\Desktop\\Practice\\2011.csv")
```

##### Assess structure 

```{r eval = TRUE}
# Ensure formats are appropriate for associated variables
# No formatting changes are necessary since chr, int, and dbl are assigned correctly.

glimpse(provider)

```

##### Explore missing values 

```{r eval = TRUE}
# No missing values were found overall, but will double check each variable individually.
sum(is.na(provider))

sum(is.na(provider$`DRG Definition`))

sum(is.na(provider$`Provider Id`))

sum(is.na(provider$`Provider Name`))

sum(is.na(provider$`Provider Street Address`))

sum(is.na(provider$`Provider City`))

sum(is.na(provider$`Provider State`))

sum(is.na(provider$`Provider Zip Code`))

sum(is.na(provider$`Hospital Referral Region Description`))

sum(is.na(provider$`Total Discharges`))

sum(is.na(provider$`Average Covered Charges`))

sum(is.na(provider$`Average Total Payments`))

sum(is.na(provider$`Average Medicare Payments`))
```

##### Separate columns

```{r eval = TRUE}
# The DRG number and the associated medical condition are under one variable.
# Ultimately want to assess if each DRG number matches a unique condition.
# It will be easier to compare DRG numbers to descriptions if the columns are separated.

provider_sepDRG <- provider %>% separate(col = `DRG Definition`,
                                         into = c('DRG Number', 'DRG Description'),
                                         sep = ' - ', remove = FALSE)
View(provider_sepDRG)
```

##### Assess DRG number vs. Definition

```{r eval = TRUE}
# Each of the 100 DRG numbers does indeed correspond to a unique medical description.
# Viewing the output in a data frame makes it easier to see the 1:1 correspondence.

gut_check <- data.frame(provider_sepDRG %>% group_by(`DRG Number`) %>% 
                          summarise("unique_descrip" = n_distinct(`DRG Description`)))
gut_check
```
##### Rename variables 

```{r eval = TRUE}
# Viewing the data in a table format via the console revealed tha all the variables are encased 
#in backticks.  Backticks can sometimes cause a problem in statistical analysis.
# The variables will be renamed to exclude the backticks.
# Will also rename key variables by adding the word "Provider" to differentiate provider
# variables from National and State

# View tbl in console to see backticks

tbl_df(provider_sepDRG) 

provider_sepDRG <- provider_sepDRG %>% 
  rename(DRG_Definition = `DRG Definition`, DRG_Number = `DRG Number`,
         DRG_Description = `DRG Description`, Provider_ID = `Provider Id`,
         Provider_Name = `Provider Name`, Provider_Street_Address = `Provider Street Address`,
         Provider_City = `Provider City`, Provider_State = `Provider State`,
         Provider_Zip_Code = `Provider Zip Code`, 
         Hospital_Referral_Region_Description = `Hospital Referral Region Description`,
         Provider_Total_Discharges = `Total Discharges`, 
         Provider_Avg_Covered_Charges = `Average Covered Charges`, 
         Provider_Avg_Total_Payments = `Average Total Payments`, 
         Provider_Avg_Medicare_Payments = `Average Medicare Payments`)

# Confirm in console that columns are renamed

tbl_df(provider_sepDRG) 

```
##### Examine variable range

```{r eval = TRUE}
# Screen int and dbl columns to identify any negative values (don't worry about zip codes)

# No negative values were found

provider_sepDRG$Provider_Total_Discharges %>% range()
provider_sepDRG$Provider_Avg_Covered_Charges %>% range()
provider_sepDRG$Provider_Avg_Total_Payments %>% range()
provider_sepDRG$Provider_Avg_Medicare_Payments %>% range()

```
##### The steps above yield a cleaned-up file called provider_sepDRG at the  hospital level.

```{r eval = TRUE}

# Save a cleaned up provider file for future data analysis.

write_csv(x = provider_sepDRG, 
          path = "C:\\Users\\bliss\\Documents\\R\\Capstone\\Medicare\\Test\\provider_sepDRG.csv")

```


# _____________________________________________________________________

### 2. Clean National and State files   
  

#### Medicare also provides 2011 National and State files.
 

##### Load national data and repeat similar cleaning exercises as in Provider file.


```{r eval = TRUE}
National_Average_CMS <- read.csv("C:\\Users\\bliss\\Desktop\\Practice\\2011_National.csv")

# Confirm structure was similar to provider file

glimpse(National_Average_CMS)

# Explore missing values - there are none

sum(is.na(National_Average_CMS))

# Separate DRG number from description
National_Average_CMS <- National_Average_CMS %>% 
  separate(col = DRG.Definition, into = c("DRG_Number", "DRG_Definition"), sep = " - ",
           remove = FALSE)

# Check there are 100 unique DRG Numbers and Definitions, as seen in the provider file

check <- data.frame(National_Average_CMS %>% group_by(DRG_Number) %>%
  summarise(n_distinct(DRG_Definition)))

check # View in console

# Rename columns so it is clear they are National ones, and also remove backticks

tbl_df(National_Average_CMS) # View in console to see backticks

colnames(National_Average_CMS)

National_Average_CMS <- National_Average_CMS %>% rename("Nat_Total_Discharges" =
                                                          Total.Discharges,
                                                        "Nat_Avg_Covered_Charges" = 
                                                          Average.Covered.Charges,
                                                        "Nat_Avg_Total_Payments" =
                                                          Average.Total.Payments,
                                                        "Nat_Avg_Medicare_Payments" =
                                                          Average.Medicare.Payments)
tbl_df(National_Average_CMS) # View in console to confirm

View(National_Average_CMS)

# Check int and dbl columns for negative values (don't worry about zip codes)

National_Average_CMS$Nat_Total_Discharges %>% range()
National_Average_CMS$Nat_Avg_Covered_Charges %>% range()
National_Average_CMS$Nat_Avg_Total_Payments %>% range()
National_Average_CMS$Nat_Avg_Medicare_Payments %>% range()

```

```{r eval = TRUE}

# Check int and dbl columns for negative values (don't worry about zip codes)

National_Average_CMS$Nat_Total_Discharges %>% range()
National_Average_CMS$Nat_Avg_Covered_Charges %>% range()
National_Average_CMS$Nat_Avg_Total_Payments %>% range()
National_Average_CMS$Nat_Avg_Medicare_Payments %>% range()

```


```{r eval = TRUE}
# Save a cleaned up version of the National file

write_csv(x = National_Average_CMS, path =
            "C:\\Users\\bliss\\Documents\\R\\Capstone\\Medicare\\Test\\National_Average_CMS.csv")

```


##### Load state data and repeat similar cleaning exercises as in Provider file.  

```{r eval = TRUE}

State_Average_CMS <- read.csv("C:\\Users\\bliss\\Desktop\\Practice\\2011_State.csv")

# Confirm structure was similar to provider file

glimpse(State_Average_CMS)

View(State_Average_CMS)

# Check for missing values - there are none

sum(is.na(State_Average_CMS))

# Separate DRG number from definition

State_Average_CMS <- State_Average_CMS %>%
  separate(col = DRG.Definition, into = c("DRG_Number", "DRG_Definition"), sep = " - ",
           remove = FALSE)

# Confirm 100 unique DRGs

check_1 <- State_Average_CMS %>% group_by(DRG_Number) %>%
  summarise(n_distinct(DRG_Definition))

data.frame(check_1)

# Rename columns so it is clear they are State ones

tbl_df(State_Average_CMS) # See if there are any backticks

colnames(State_Average_CMS)

State_Average_CMS <- State_Average_CMS %>% rename("State_Total_Discharges" =
                                                          Total.Discharges,
                                                        "State_Avg_Covered_Charges" = 
                                                          Average.Covered.Charges,
                                                        "State_Avg_Total_Payments" =
                                                          Average.Total.Payments,
                                                        "State_Avg_Medicare_Payments" =
                                                          Average.Medicare.Payments,
                                                  "State_Provider" = Provider.State)

View(State_Average_CMS)

```

```{r eval = TRUE}

# Check int and dbl columns for negative values (don't worry about zip codes)

State_Average_CMS$State_Total_Discharges %>% range()
State_Average_CMS$State_Avg_Covered_Charges %>% range()
State_Average_CMS$State_Avg_Total_Payments %>% range()
State_Average_CMS$State_Avg_Medicare_Payments %>% range()

```

```{r eval = TRUE}

# Save a cleaned up version of the State file

write_csv(x = State_Average_CMS, path =
            "C:\\Users\\bliss\\Documents\\R\\Capstone\\Medicare\\Test\\State_Average_CMS.csv")

```

### 3. Merge the National file with the Provider file by DRG number

```{r eval = TRUE}

# Merge National file 

provider_sepDRG_Nat <- left_join(x = provider_sepDRG, y = National_Average_CMS,
                                       by = "DRG_Number")

# Remove duplicate column names related to DRG information

colnames(provider_sepDRG_Nat)

provider_sepDRG_Nat <- provider_sepDRG_Nat %>% select(-c(DRG.Definition, DRG_Definition.y))

View(provider_sepDRG_Nat)

colnames(provider_sepDRG_Nat)

# double check to make sure there are 100 unique National DRG Covered charges in the merged file

check <- data.frame(provider_sepDRG_Nat %>% group_by(DRG_Number) %>% 
                      summarise(National = n_distinct(Nat_Avg_Covered_Charges)))

check # View in console


```

```{r}
# Save merged Provider_National file

write_csv(x = provider_sepDRG_Nat, path =
            "C:\\Users\\bliss\\Documents\\R\\Capstone\\Medicare\\Test\\provider_sepDRG_Nat.csv")
```

##### Now we have a saved, merged file called provider_sepDRG_Nat.

### The End


