---
title: "2011_Medicare_Cleanup"
output: html_document
---
# Test

### The first objective is to clean up the 2011 Medicare hospital provider file for analysis.
* provider_sepDRG

### The second objective is to merge the National and State files with the provider file.
* provider_sepDRG_Nat_State


### 1. Clean Provider File
\  

##### Load package 
```{r eval = FALSE}
library(tidyverse)
```

##### Load provider data 

```{r eval = FALSE}
# Do I need to use setwd() so this file can be opened up from GitHub?
# This file contains data at the individual hopsital (provider) level.

provider <- read_csv("C:\\Users\\bliss\\Desktop\\Practice\\2011.csv")
```

##### Assess structure 

```{r eval = FALSE}
# Ensure formats are appropriate for associated variables
# No formatting changes are necessary since chr, int, and dbl are assigned correctly.

glimpse(provider)

```

##### Explore missing values 

```{r eval = FALSE}
# No missing values were found overall, but will double check each variable individually.
sum(is.na(provider))

sum(is.na(provider$`DRG Definition`))

sum(is.na(provider$`Provider Id`))

sum(is.na(provider$`Provider Name`))

sum(is.na(provider$`Provider Street Address`))

sum(is.na(provider$`Provider City`))

sum(is.na(provider$`Provider State`))

sum(is.na(provider$`Provider Zip Code`))

sum(is.na(provider$`Hospital Referral Region Description`))

sum(is.na(provider$`Total Discharges`))

sum(is.na(provider$`Average Covered Charges`))

sum(is.na(provider$`Average Total Payments`))

sum(is.na(provider$`Average Medicare Payments`))
```

##### Separate columns

```{r eval = FALSE}
# The DRG number and the associated medical condition are under one variable.
# Ultimately want to assess if each DRG number matches a unique condition.
# It will be easier to compare DRG numbers to descriptions if the columns are separated.

provider_sepDRG <- provider %>% separate(col = `DRG Definition`,
                                         into = c('DRG Number', 'DRG Description'),
                                         sep = ' - ', remove = FALSE)
View(provider_sepDRG)
```

##### Assess DRG number vs. Definition

```{r eval = FALSE}
# Each of the 100 DRG numbers does indeed correspond to a unique medical description.
# Viewing the output in a data frame makes it easier to see the 1:1 correspondence.

gut_check <- data.frame(provider_sepDRG %>% group_by(`DRG Number`) %>% 
                          summarise("unique_descrip" = n_distinct(`DRG Description`)))
gut_check
```
##### Rename variables 

```{r eval = FALSE}
# Viewing the data in a table format via the console revealed tha all the variables are encased #in backticks.  Backticks can sometimes cause a problem in statistical analysis.
# The variables will be renamed to exclude the backticks.
# Will also rename key variables by adding the word "Provider" to differentiate provider
# variables from National and State

# View tbl in console to see backticks

tbl_df(provider_sepDRG) 

provider_sepDRG <- provider_sepDRG %>% 
  rename(DRG_Definition = `DRG Definition`, DRG_Number = `DRG Number`,
         DRG_Description = `DRG Description`, Provider_ID = `Provider Id`,
         Provider_Name = `Provider Name`, Provider_Street_Address = `Provider Street Address`,
         Provider_City = `Provider City`, Provider_State = `Provider State`,
         Provider_Zip_Code = `Provider Zip Code`, 
         Hospital_Referral_Region_Description = `Hospital Referral Region Description`,
         Provider_Total_Discharges = `Total Discharges`, Provider_Avg_Covered_Charges = `Average Covered Charges`, Provider_Avg_Total_Payments = `Average Total Payments`, Provider_Avg_Medicare_Payments = `Average Medicare Payments`)

# Confirm in console that columns are renamed

tbl_df(provider_sepDRG) 

```
##### Examine variable range

```{r eval = FALSE}
# Screen int and dbl columns to identify any negative values (don't worry about zip codes)

# No negative values were found

provider_sepDRG$Provider_Total_Discharges %>% range()
provider_sepDRG$Provider_Avg_Covered_Charges %>% range()
provider_sepDRG$Provider_Avg_Total_Payments %>% range()
provider_sepDRG$Provider_Avg_Medicare_Payments %>% range()

```
##### The steps above yield a cleaned-up file called provider_sepDRG at the  hospital level.

# ----------------------------------------------------------------------------------

### 2. Merge National and State files with Provider  
\  

#### Medicare also provides 2011 National and State files.
\  

##### Load national data and repeat similar cleaning exercises as in provider file.

```{r eval = FALSE}
National_Average_CMS <- read.csv("C:\\Users\\bliss\\Desktop\\Practice\\2011_National.csv")

# Confirm structure was similar to provider file

glimpse(National_Average_CMS)

# Explore missing values - there are none

sum(is.na(National_Average_CMS))

# Separate DRG number from description
National_Average_CMS <- National_Average_CMS %>% 
  separate(col = DRG.Definition, into = c("DRG_Number", "DRG_Definition"), sep = " - ",
           remove = FALSE)

# Check there are 100 unique DRG Numbers and Definitions, as seen in the provider file

check <- data.frame(National_Average_CMS %>% group_by(DRG_Number) %>%
  summarise(n_distinct(DRG_Definition)))

check # View in console

# Rename columns so it is clear they are National ones, and also remove backticks

tbl_df(National_Average_CMS) # View in console to see backticks

colnames(National_Average_CMS)

National_Average_CMS <- National_Average_CMS %>% rename("Nat_Total_Discharges" =
                                                          Total.Discharges,
                                                        "Nat_Avg_Covered_Charges" = 
                                                          Average.Covered.Charges,
                                                        "Nat_Avg_Total_Payments" =
                                                          Average.Total.Payments,
                                                        "Nat_Avg_Medicare_Payments" =
                                                          Average.Medicare.Payments)
tbl_df(National_Average_CMS) # View in console to confirm

View(National_Average_CMS)

# Check int and dbl columns for negative values (don't worry about zip codes)

National_Average_CMS$Nat_Total_Discharges %>% range()
National_Average_CMS$Nat_Avg_Covered_Charges %>% range()
National_Average_CMS$Nat_Avg_Total_Payments %>% range()
National_Average_CMS$Nat_Avg_Medicare_Payments %>% range()

```
##### Load state data and repeat similar cleaning exercises as in provider file.  

```{r eval = FALSE}

State_Average_CMS <- read.csv("C:\\Users\\bliss\\Desktop\\Practice\\2011_State.csv")

# Confirm structure was similar to provider file

glimpse(State_Average_CMS)

View(State_Average_CMS)

# Check for missing values - there are none

sum(is.na(State_Average_CMS))

# Separate DRG number from definition

State_Average_CMS <- State_Average_CMS %>%
  separate(col = DRG.Definition, into = c("DRG_Number", "DRG_Definition"), sep = " - ",
           remove = FALSE)

# Confirm 100 unique DRGs

check_1 <- State_Average_CMS %>% group_by(DRG_Number) %>%
  summarise(n_distinct(DRG_Definition))

data.frame(check_1)

# Rename columns so it is clear they are State ones

tbl_df(State_Average_CMS) # See if there are any backticks

colnames(State_Average_CMS)

State_Average_CMS <- State_Average_CMS %>% rename("State_Total_Discharges" =
                                                          Total.Discharges,
                                                        "State_Avg_Covered_Charges" = 
                                                          Average.Covered.Charges,
                                                        "State_Avg_Total_Payments" =
                                                          Average.Total.Payments,
                                                        "State_Avg_Medicare_Payments" =
                                                          Average.Medicare.Payments,
                                                  "State_Provider" = Provider.State)

View(State_Average_CMS)

# Check int and dbl columns for negative values (don't worry about zip codes)

State_Average_CMS$State_Total_Discharges %>% range()
State_Average_CMS$State_Avg_Covered_Charges %>% range()
State_Average_CMS$State_Avg_Total_Payments %>% range()
State_Average_CMS$State_Avg_Medicare_Payments %>% range()
```
##### Merge national and state file with provider file by the DRG number.

```{r eval = FALSE}

# Merge national file

provider_sepDRG_Nat_State <- left_join(x = provider_sepDRG, y = National_Average_CMS,
                                       by = "DRG_Number")

colnames(provider_sepDRG_Nat_State)

# Remove repeat columns that are not needed

provider_sepDRG_Nat_State <- provider_sepDRG_Nat_State %>% select (-c(DRG.Definition, DRG_Definition.y, DRG_Definition.y))

View(provider_sepDRG_Nat_State)

# Join State table and remove repeat columns.

provider_sepDRG_Nat_State <- left_join(x = provider_sepDRG_Nat_State, y = State_Average_CMS,
                                       by = "DRG_Number")

colnames(provider_sepDRG_Nat_State)

provider_sepDRG_Nat_State <- provider_sepDRG_Nat_State %>% select(-c(State_Provider,
                                                                     DRG_Definition,
                                                                     DRG.Definition))

colnames(provider_sepDRG_Nat_State)

# Check to make sure there are 100 unique National DRG Covered Charges

check <- data.frame(provider_sepDRG_Nat_State %>% group_by(DRG_Number) %>% summarise(National = 
                                                                   n_distinct(Nat_Avg_Covered_Charges)))

check # View in console

# Check to make sure there are 51 unique State DRG Covered Charges (includes Washington DC)

check_1 <- data.frame(original_sepDRG_Nat_State %>% group_by(DRG_Number) %>%
  summarise(State = n_distinct(State_Avg_Covered_Charges)))

check_1 # View in console
```
##### Now we have a merged file called provider_sepDRG_Nat_State.
